---
title: Finnish MPs on Twitter part 1
date: '2020-07-30'
slug: finnish-mps-on-twitter-part1
categories:
  - '#twitter #nlp'
tags: []
type: ''
subtitle: ''
image: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

A great qualitative analysis of language used by Finnish members of the parliament (MPs) on Twitter was made couple days ago (in Finnish). 

```{r echo=FALSE}
blogdown::shortcode('tweet', '1287726197607149568')
```


Here is my take on analysis of tweets and retweets of Finnish MPs. Goals for this project:

* Automate hourly collection of tweet data produced by MPs.
* Describe how Finnish MPs tweet based on their background variables. At least by party, age, gender and mp experience years.
* Sentiment analysis of the language used in the tweets. 
* Update constantly open source raw data and publish weekly a report on language used.

Of course all this done when 👶 allows me to code. 

## Current Finnish MPs

First I scrape the listing of current MPs and their backgrounds from Wikipedia https://fi.wikipedia.org/wiki/Luettelo_vaalikauden_2019%E2%80%932023_kansanedustajista. Package `rvest` is used here.

```{r message=FALSE}
library(rvest)
library(tidyverse)
library(janitor)
url <- "https://fi.wikipedia.org/wiki/Luettelo_vaalikauden_2019%E2%80%932023_kansanedustajista"

wikipedia <- read_html(url)

seniority <- function(period) {
  start <- period %>%
    str_extract("\\d{4}(?=–)")
  end <- period %>%
    str_extract("(?<=–)\\d{4}") 
  if(is.na(end)) return(1)
  start:end %>% 
    length()

}

wikipedia <- wikipedia %>% 
  html_nodes("table") %>% 
  .[[2]] %>% 
  html_table() %>% 
  clean_names %>% 
  filter(puolue == "") %>% 
  mutate(party = str_remove(puolue_2, "\\[.*\\]")) %>% 
  select(-puolue_2,
         -puolue,
         xp_in_years = kansanedustajana) %>% 
  mutate(xp_in_years = str_split(xp_in_years, ", ")) %>% 
  unnest(xp_in_years) %>% 
  rowwise() %>% 
  mutate(xp_in_years = seniority(xp_in_years)) %>% 
  group_by(nimi) %>% 
  mutate(xp_in_years = sum(xp_in_years)) %>% 
  ungroup() %>% 
  distinct %>% 
  mutate(join_nimi = nimi %>% 
           str_to_lower %>% 
           str_remove(" |\\t|\\s") %>% 
           stringi::stri_trans_general("Latin-ASCII")) %>% 
  rename(electoral_district = vaalipiiri,
         mp_name = nimi)
```

## Current Tweeting Finnish MPs

I was thankfully notified that Finnish Parliament maintains a list of the MPs who are on Twitter.

```{r echo=FALSE}
blogdown::shortcode('tweet', '1287781573308420096')
```

Lets start collecting all accounts from that list with `rtweet` package.

```{r message=FALSE}
library(rtweet)

twiittaavat_kansanedustajat <- lists_members("203337069") %>% 
  select(matches("_id|name|_count$|at$"))
```

Some MP Twitter account names have nicknames, emojis, Arabic and typos and therefore combining of the Twitter and Wikipedia needs a small manual mapping for some MP names.

```{r message=FALSE}
twiittaavat_kansanedustajat <- twiittaavat_kansanedustajat %>%
  mutate(name = case_when(
    str_detect(name, "Harkimo") ~ "Harry Harkimo",
    str_detect(name, "Ari Torniainen") ~ "Ari Torniainen",
    str_detect(name, "Taee") ~ "Hussein al-Taee",
    str_detect(name, "Eeva-Johanna Elorant") ~ "Eeva-Johanna Eloranta",
    str_detect(name, "Niina Malm") ~ "Niina Malm",
    str_detect(name, "Johanna Ojala") ~ "Johanna Ojala-Niemelä",
    str_detect(name, "Petelius") ~ "Pirkka-Pekka Petelius",
    str_detect(name, "Mäkisalo-R") ~ "Merja Mäkisalo-Ropponen",
    TRUE ~ name),
    join_nimi = name %>% 
      str_to_lower %>% 
      str_remove(" |\\t|\\s") %>% 
      stringi::stri_trans_general("Latin-ASCII")) %>%
  rename(twitter_name = name) %>% 
  left_join(wikipedia) %>% 
  select(-join_nimi)
```

And voila. Here we have tweeting MPs, their twitter metadata, party affiliation, MP experience in years and electoral district.

```{r echo=FALSE}
glimpse(twiittaavat_kansanedustajat) 
```
## Collect Status dataset

Median MP has ~1800 statuses. Status is a tweet or a retweet.  Twitter API allows collecting last 3200 status messages. That is an amount that 60 percent of MPs have during their whole history. 

```{r}
twiittaavat_kansanedustajat %>% 
  ggplot() + aes(statuses_count) + geom_density() + geom_vline(xintercept = median(twiittaavat_kansanedustajat$statuses_count))
```

Collecting maximum of 3200 statuses per MP sums to total of 310 000 tweets.

```{r echo=FALSE}

# statuses <- twiittaavat_kansanedustajat$user_id %>% 
#   map(function(x){
#     gc()
#     Sys.sleep(7)
#     get_timeline(x, n = 3200)
#   })
# 

# statuses <- statuses %>%
#   bind_rows()
# 
# statuses %>% 
#   saveRDS("data/statuses.rds")

statuses <- readRDS("data/statuses.rds")

glimpse(statuses)
```

MPs produce weekly thousands of statuses. Sometimes over 3000 statuses.

```{r}
statuses %>% 
  mutate(year = lubridate::year(created_at),
         week = lubridate::week(created_at)) %>%
  filter(year > 2018) %>% 
  count(year, week, name = "weekly_statuses") %>% 
  ggplot + aes(weekly_statuses) + geom_density()
```


## Next up

Next time I will classify statuses by their sentiments.
